### 2024-03-10 chap_atom_mol/chap_atom_mol.tex
Edit since UQ_TD_full has been completed

@@

\begin{table}[htbp!]
\caption{Total Sobol' indices of the uncertain parameters over the total transition cost. Where the cost of purchasing electrofuels is the top-1 parameter, \gls{SMR}-related parameters have a negligigble impact on this cost.}
\label{tab:UQ_short}
\centering
\begin{tabular}{l c c}
\toprule
\textbf{Parameter}  & \textbf{Ranking} & \textbf{Sobol' index} \\	
\midrule
\textbf{Purchase electrofuels} & 1 & 47.4\% \\
Industry EUD & 2 & 23.5\% \\
Interest rate & 3 & 11.0\% \\
Purchase fossil fuels  & 4 & 6.9\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{Potential capacity \gls{SMR}} & 11 & 0.9\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{CAPEX \gls{SMR}} & 32 & $<$0.1\% \\
\bottomrule

@@

Given the relatively wide uncertainty range (\ie up to [-30.8\%; +24.0\%] by 2050) and, above all, the major share among the total demand, between 53\% and 60\%, the industrial \gls{EUD} is the second most impacting parameter. Then, as the driving factor for the annualisation and the salvage value of the assets, the interest rate has a 11\% Sobol' index\footnote{It is important to note here that the model considers an overall interest rate for the entire system (\ie 1.5\% as a nominal value). In practice, the interest rate would vary depending on the technology investment risk. This variation would have, for instance, a major impact on the \gls{LCOE} of technologies like nuclear power plants \cite{world_nuclear_asso}, given the important capital needs and long time horizons \cite{IEA_Nuclear_2022}.}. Finally, similarly to electrofuels, the cost of purchasing fossil fuels is also to consider in the perspective to reduce the uncertainty over the total transition cost. However, due to the ambitious \ce{CO2}-budget, phasing out of fossil fuels is urgent and makes their uncertain impact smaller than their renewable alternatives.

\Cref{fig:UQ_PDF_total_transition_cost} shows the \gls{PDF} of the total transition cost given the 1260 samples. Stretching between 680\,b€ and 1940\,b€, the mean, the median and the nominal value (\ie REF case) are close to each other, respectively 1160\,b€, 1150\,b€ and 1080\,b€. Similarly to the analysis carried out by \citet{coppitters2023optimizing}, one observes here that the distribution is right-skewed. It could then be qualified as ``fragile" as the top 50\% of the samples cover a bigger range (\ie between 1150 and 1940) than the bottom 50\% of the samples (\ie between 680 and 1150). In other words, the bad scenarios, resulting in a total transition cost higher than the median, have a bigger effect on this cost.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_PDF_total_transition_cost.pdf}
\caption{\acrfull{PDF} of the total transition cost. The mean, $\mu=1.16\cdot10^3$\,b€, is slightly higher than the median ($P_{50}=1.15\cdot10^3$\,b€) and the nominal cases cost, $1.08\cdot10^3$\,b€ and $1.04\cdot10^3$\,b€ respectively for the REF and SMR cases. Also, with a standard deviation, $\sigma=185$\,b€, a 95\%-confidence interval would be about [0.8; 1.5]$\cdot10^3$\,b€.}
\label{fig:UQ_PDF_total_transition_cost}
\end{figure}

@@

As the general trends are increasing, discrepancies exist between the different energy carriers. E-methane, as the renewable alternative to fossil natural gas, substitutes it, sometimes at a very early stage of the transition, 2025, and to a somehow unrealistically large extent, 173\,TWh, which is more than 13\% more than the total import of electrofuels in the REF case. The necessity to import this molecule is progressive through the transition to supply mostly industrial \gls{CHP} and boilers.

E-hydrogen becomes rapidly the main stream of hydrogen in the system, on top of steam-methane-reforming or electrolysis, to reach a median and a maximum values of, respectively 13.6\,TWh and 40.7\,TWh in 2050. Hydrogen is more frequently used in the mobility sector. Like in the REF case, fuel cell trucks are often the first option but, in some outlying cases, fuel cell cars and buses appear to completely substitute respectively \gls{BEV} cars and \gls{CNG} buses by 2050. Moreover, some samples lead to local production of methanol via the methanolation process, to produce up to 16.3\,TWh of methanol (\ie 30\% of the total supply of methanol of the nominal REF and SMR cases).

Then, the imported e-ammonia, becoming rapidly cost-competitive against its fossil alternative (\Cref{fig:cs_resources_cost}), quickly substitutes it and the Haber-Bosch process. Where the initial purpose of ammonia is to satisfy a relatively limited \acrfull{NED} (\ie 10 $\pm$ 3\,TWh by 2050), the variation of its import is mostly due to the higher or lower need for ammonia-\gls{CCGT} as a flexible option to produce electricity. From 2035, out of the four considered electrofuels, the imported e-ammonia is the one exhibiting the largest uncertainty with, for instance, an interquartile range (IQR)\footnote{The interquartile range is the difference between the third quartile ($Q3$ or $P_{75}$) and the first ($Q1$ or $P_{25}$). It is an indicator of statistical dispersion around the median, $Q2$ or $P_{50}$.} of about 50\,TWh. In some extreme cases, e-ammonia is the most imported molecules, \ie up to 233\,TWh or 63\% of the total primary mix in the REF case in 2050.

>>>>

Beyond the direct impact that \gls{SMR} has on the power sector, it is worth assessing the other sectors given the sector-coupling effect related to the whole-energy system optimisation \cite{contino2020whole}.

@@

\begin{table}[htbp!]
\caption{Total Sobol' indices of the uncertain parameters over the total transition cost. Where the cost of purchasing electrofuels is the top-1 parameter, \gls{SMR}-related parameters have a negligigble impact on this cost.}
\label{tab:UQ_short}
\centering
\begin{tabular}{l c c}
\toprule
\textbf{Parameter}  & \textbf{Ranking} & \textbf{Sobol' index} \\	
\midrule
\textbf{Purchase electrofuels} & 1 & 46.8\% \\
Industry EUD & 2 & 23.2\% \\
Interest rate & 3 & 12.0\% \\
Purchase fossil fuels  & 4 & 5.7\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{Potential capacity \gls{SMR}} & 11 & 0.9\% \\
$\vdots$ & $\vdots$ & $\vdots$\\
\textbf{CAPEX \gls{SMR}} & 33 & $<$0.1\% \\
\bottomrule

@@

Given the relatively wide uncertainty range (\ie up to [-30.8\%; +24.0\%] by 2050) and, above all, the major share among the total demand, between 53\% and 60\%, the industrial \gls{EUD} is the second most impacting parameter. Then, as the driving factor for the annualisation and the salvage value of the assets, the interest rate has a 12\% Sobol' index\footnote{It is important to note here that the model considers an overall interest rate for the entire system (\ie 1.5\% as a nominal value). In practice, the interest rate would vary depending on the technology investment risk. This variation would have, for instance, a major impact on the \gls{LCOE} of technologies like nuclear power plants \cite{world_nuclear_asso}, given the important capital needs and long time horizons \cite{IEA_Nuclear_2022}.}. Finally, similarly to electrofuels, the cost of purchasing fossil fuels is also to consider in the perspective to reduce the uncertainty over the total transition cost. However, due to the ambitious \ce{CO2}-budget, phasing out of fossil fuels is urgent and makes their uncertain impact smaller than their renewable alternatives.

\Cref{fig:UQ_PDF_total_transition_cost} shows the \gls{PDF} of the total transition cost given the 1260 samples. Stretching between 660\,b€ and 2050\,b€, the mean, the median and the nominal value (\ie REF case) are close to each other, respectively 1180\,b€, 1160\,b€ and 1080\,b€. Similarly to the analysis carried out by \citet{coppitters2023optimizing}, one observes here that the distribution is right-skewed. It could then be qualified as ``fragile" as the top 50\% of the samples cover a bigger range (\ie between 1160 and 2050) than the bottom 50\% of the samples (\ie between 660 and 1160). In other words, the bad scenarios, resulting in a total transition cost higher than the median, have a bigger effect on this cost.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_PDF_total_transition_cost.pdf}
\caption{\acrfull{PDF} of the total transition cost. The mean, $\mu=1.18\cdot10^3$\,b€, is slightly higher than the median ($P_{50}=1.16\cdot10^3$\,b€) and the nominal cases cost, $1.08\cdot10^3$\,b€ and $1.04\cdot10^3$\,b€ respectively for the REF and SMR cases. Also, with a standard deviation, $\sigma=197$\,b€, a 95\%-confidence interval would be about [0.8; 1.6]$\cdot10^3$\,b€.}
\label{fig:UQ_PDF_total_transition_cost}
\end{figure}

@@

As the general trends are increasing, discrepancies exist between the different energy carriers. E-methane, as the renewable alternative to fossil natural gas, substitutes it, sometimes at a very early stage of the transition, 2025, and to a somehow unrealistically large extent, 163\,TWh, which is more than 6\% more than the total import of electrofuels in the REF case. The necessity to import this molecule is progressive through the transition to supply mostly industrial \gls{CHP} and boilers.

E-hydrogen becomes rapidly the main stream of hydrogen in the system, on top of steam-methane-reforming or electrolysis, to reach a median and a maximum values of, respectively 13.0\,TWh and 42.1\,TWh in 2050. Hydrogen is more frequently used in the mobility sector. Like in the REF case, fuel cell trucks are often the first option but, in some outlying cases, fuel cell cars and buses appear to completely substitute respectively \gls{BEV} cars and \gls{CNG} buses by 2050. Moreover, some samples lead to local production of methanol via the methanolation process, to produce up to 17.8\,TWh of methanol (\ie 33\% of the total supply of methanol of the nominal REF and SMR cases).

Then, the imported e-ammonia, becoming rapidly cost-competitive against its fossil alternative (\Cref{fig:cs_resources_cost}), quickly substitutes it and the Haber-Bosch process. Where the initial purpose of ammonia is to satisfy a relatively limited \acrfull{NED} (\ie 10 $\pm$ 3\,TWh by 2050), the variation of its import is mostly due to the higher or lower need for ammonia-\gls{CCGT} as a flexible option to produce electricity. From 2035, out of the four considered electrofuels, the imported e-ammonia is the one exhibiting the largest uncertainty with, for instance, an interquartile range (IQR)\footnote{The interquartile range is the difference between the third quartile ($Q3$ or $P_{75}$) and the first ($Q1$ or $P_{25}$). It is an indicator of statistical dispersion around the median, $Q2$ or $P_{50}$.} of about 50\,TWh. In some extreme cases, e-ammonia is the most imported molecules, \ie up to 216\,TWh or 58\% of the total primary mix in the REF case in 2050.

### 2024-03-10 chap_intro_ccl/conclusion.tex
Merge taxonomy paper + Implement Hervé's comments on case_study
Edit since UQ_TD_full has been completed

>>>>

\begin{itemize}
\item \textbf{Word about sufficiency} oui mais c'est sans doute celle qui est le plus enviable car la moins risqué à tenter d'atteindre, les solutions mirages technologiques, si on y croit trop on met tout la dessus et si ca foire, on est encore plus dans la merde pcq la direction est mauvaise, la solution sobriété, meme si jamais atteint les efforts pr l'atteindre ne seront pas contre productif
\item \textbf{Word about availability of electrfuels} Voir mémoire Ced et Simon
\item Extract a roadmap representative of a multiple-run UQ or RL process.
\end{itemize}
